\documentclass[12pt, oneside]{memoir}

\title{The Ronald Reagan Library Experience}
\author{Walt Askew}
\date{\today}

\begin{document}
\maketitle

Lines have an inhuman quality.
There's a harsh difference between sitting in a moving subway car
and one that's halted for subway car reasons.
Nothing about the actual experience of sitting in the car has changed.
But as soon as motion stops, passengers start exchanging looks of
disgust and personal injury.
Rolling eyes.
Exasperated sighs.
A particular spot at the base of the skull begins to itch and grumble
and transform into the physical site of cynicism about ``tax dollars at
work.''

I involuntarily start to scratch at that spot in the brain where
conservative politics originates as I wait.
The slapping tacking noise of colliding plastic lips falls over the
animatron's words, 
``I think you all know that I've always felt the nine most terrifying
words in the English language are: I'm from the Government, and I'm
here to help.''

``Damn straight!'' shouts the man behind me.
He high fives his friend.

The line finally pulls us towards the roller coaster.
An attendant in a bright red sweater with a plastic American flag pin
and a nest of white hair smiles as she buckles me in.
``Enjoy yourself, hon!''
I nod.

The excitement of the two friends behind me shakes forward into my car.
``I'm going to keep my arms up the entire time!''  

``No way --- no way you even make it through Iran-Contra!''

``The whole time, man. THE WHOLE TIME, for sure!''

The anxious clacking of the roller coaster towing up towards
inevitability sounds over a popping recording,
``Our government has a firm policy not to capitulate to terrorist
demands. We did not, repeat, did not trade weapons or anything else
for hostages. Nor will we.''

``My brother totally shit himself on this ride!''

``This is going to be epic!''

We crest the hill and my heart pulls into my throat which wants to
vomit it out with every bit of patriotism forcing its way into my
bloodstream.

``A few months ago, I told the American people I did not trade arms
for hostages. My heart and my best intentions still tell me that's
true, but the facts and the evidence tell me it is not''

The coaster rushed through valve's of the Gipper's enormous heart.
I learned it was full of tax cuts and fiscal responsibility.
It spiraled down and pulled us sideways in a dance and a gamble with
gravity.
I turned my head towards the bottom of the spiral and budget cuts
stared back into the center of my vision.
They were stern. Yet caring.
``There's no way! No way! This is even better than the Jeff Davis
waterpark!'' 

``General Secretary Gorbachev, if you seek peace, if you seek
prosperity for the Soviet Union and Eastern Europe, if you seek
liberalization, come here to this gate.  Mr. Gorbachev, open this
gate! Mr. Gorbachev, tear down this wall!''

The coaster ran at full speed towards the Berlin Wall which, at the
last possible moment, broke before the coaster as it blasted through.

``Holy shit, dude! Holy shit! That was so close!''

``I thought we were going to splattered!''

``Whoo! Yeah! Fuck that wall!''

``Get fucked, Gorbachev!''

``Glasnost and perestroika my ass!''

A woman bearing the pain of a nine-to-five forced smile unlocked my
harness and opened my compartment's door.
``Thank you for riding The Ronald Reagan Library
Experience\texttrademark.''
I could hear the sound of the trademark symbol in her voice.
``We hope you'll join us for another administration soon!''




% ``What began as a strategic opening to Iran deteriorated, in its.
% implementation, into trading arms for hostages.''
% 
% ``A few months ago I told the American people I did not trade arms for
% hostages. My heart and my best intentions still tell me that's true,
% but the facts and the evidence tell me it is not. As the Tower board
% reported, what began as a strategic opening to Iran deteriorated, in
% its implementation, into trading arms for hostages. This runs counter
% to my own beliefs, to administration policy, and to the original
% strategy we had in mind.''

\end{document}
